---
title: "Naxos - Exploratory Analyses"
author: "Kate Culhane"
date: "2/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}

# Load necessary packages

library(tidyverse)
library(vegan)

# Load necessary data

diet <- read_csv('./raw_data/diet_field.csv') %>% # diet data from field
  filter(island == 'Naxos')

inverts_meta <- read_csv('./raw_data/inverts_meta.csv') %>% # metadata for invert community field collection
  filter(island == 'Naxos')

pitfall <- read_csv('./raw_data/inverts_pitfall.csv') %>% # invert community data from pitfall traps
  filter(island == 'Naxos')

sticky <- read_csv('./raw_data/inverts_sticky.csv') %>% # invert community data from sticky traps
  filter(island == 'Naxos')

site_meta <- read_csv('./raw_data/site_meta.csv') # metadata for Naxos sites

site_veg <- read_csv('./raw_data/site_veg.csv') # vegetation quadrat data for Naxos sites

```


### Guiding Questions

**1: Understanding variation in the invertebrate community on Naxos**

- Is there variation in the invertebrate community across nine different sites on Naxos?
- Is that variation related to vegetation characteristics? (% cover of different proportions of vegetation)
- Is that variation related to human land-use metrics? (Presence/absence of goats or walls)

**2: Understanding diet of *P. erhardii* on Naxos**

- Is there variation in diet of the lizards across the nine sites?
- Is any variation in diet related to the variation in insect availability? (Are the lizards generalists or specialists? Are they choosy? Are they eating just in proportion to what they're seeing?)


### Analyses to do

1. NMDS for prey communities
2. NMDS for lizard diet
3. Adonis between site types