---
title: "Data Wrangling"
author: "Kate Culhane"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(include=FALSE)
```

<br>

### Pre-analysis manipulation of datasets

CSVs created for use in analyses, stored in "output_data".

<br>

**Prey data:**

- *prey.csv*: prey data from both sticky and pitfall traps, one row per trap pair
- *prey_effort.csv*: total trapping effort at each site for pitfall and sticky traps (max = 24 = 12 sticky + 12 pitfall), one row per site
- *naxos_prey.csv*: prey data from both sticky and pitfall traps, for Naxos only, one row per trap pair

<br>

**Diet data:**

- *diet.csv*: diet data from field samples, one row per lizard
- *diet_effort.csv*: total number of lizards pumped at each site, one row per site
- *naxos_diet.csv*: diet data from field samples, for Naxos only, one row per lizard

<br>

**Site by OTU matrices:**

- *naxos_prey_matrix.csv*: Site by OTU matrix for prey data, for Naxos data only
- *naxos_diet_matrix.csv*: Site by OTU matrix for diet data, for Naxos data only

<br>

```{r packages and data}

# Load packages
library(kableExtra)
library(tidyverse)

# Diet data
field_diet <- read_csv('raw_data/diet_field.csv') # from field samples

# Prey data
pitfall <- read_csv('raw_data/prey_pitfall.csv') %>% # from pitfall traps
  mutate(trap_id = paste0(site, '-', trap),
         type = 'pitfall')
sticky <- read_csv('raw_data/prey_sticky.csv') %>% # from sticky traps
  mutate(trap_id = paste0(site, '-', trap),
         type = 'sticky')
prey_meta <- read_csv('raw_data/prey_meta.csv') # trap metadata

# Naxos site data
naxos_meta <- read_csv('raw_data/naxos_meta.csv') # metadata for Naxos sites
naxos_veg <- read_csv('raw_data/naxos_veg.csv') # veg quadrat data for Naxos sites

```

```{r}

# Prey data

# Combine pitfall and sticky data by trap
prey <- bind_rows(pitfall, sticky) %>% 
  select(-c(date_analyzed, condition, comments, type)) %>% 
  group_by(island, site, trap, trap_id) %>% 
  summarize_all(sum, na.rm=TRUE) %>% 
  ungroup()
write_csv(prey, 'output_data/prey.csv', col_names=TRUE)

# Trapping effort
prey_effort <- bind_rows(pitfall, sticky) %>% 
  select(island, site, trap, trap_id, type, condition) %>% 
  mutate(
    sticky_effort = case_when(
      type=='sticky' & condition=='GOOD' ~ 1,
      type=='sticky' & condition=='FLAW' ~ 0.5,
      TRUE ~ 0
    ),
    pitfall_effort = case_when(
      type=='pitfall' & condition=='GOOD' ~ 1,
      type=='pitfall' & condition=='FLAW' ~ 0.5,
      TRUE ~ 0
    )
  ) %>% 
  group_by(island, site) %>% 
  summarize_at(c('sticky_effort', 'pitfall_effort'), sum) %>% 
  mutate(effort = (sticky_effort + pitfall_effort))
write_csv(prey_effort, 'output_data/prey_effort.csv', col_names=TRUE)

# Naxos data only
naxos_prey <- prey %>% 
  filter(island == 'Naxos') %>% 
  full_join(naxos_meta[c(1:4)]) %>% 
  select(site, use, veg, trap, trap_id, coleoptera:homoptera)
write_csv(naxos_prey, 'output_data/naxos_prey.csv', col_names=TRUE)

```

```{r}

# Diet data

# All data
diet <- field_diet %>% 
  select(-date)
write_csv(diet, 'output_data/diet.csv', col_names=TRUE)

# Trapping effort (number of lizards pumped)
diet_effort <- diet %>% 
  group_by(island, site) %>% 
  tally(name = 'n_lizards')
write_csv(diet_effort, 'output_data/diet_effort.csv', col_names=TRUE)

# Naxos data only
naxos_diet <- diet %>% 
  filter(island == 'Naxos') %>% 
  full_join(naxos_meta[c(1:4)]) %>% 
  select(site, use, veg, lizard_id, blattodea:salt, comments)
write_csv(naxos_diet, 'output_data/naxos_diet.csv', col_names=TRUE)


```

```{r}

# Site by OTU matrices

# Prey
naxos_prey_matrix <- naxos_prey %>%
  select(-c(trap, trap_id)) %>% 
  group_by(site, use, veg) %>% 
  summarize_all(sum) %>% 
  arrange(use, veg)
naxos_prey_matrix <- naxos_prey_matrix[, colSums(naxos_prey_matrix != 0) > 0]
write_csv(naxos_prey_matrix, 'output_data/naxos_prey_matrix.csv', col_names=TRUE)

# Diet
naxos_diet_matrix <- naxos_diet %>% 
  select(-c(lizard_id, comments)) %>% 
  group_by(site, use, veg) %>% 
  summarize_all(sum) %>% 
  arrange(use, veg)
naxos_diet_matrix <- naxos_diet_matrix[, colSums(naxos_diet_matrix != 0) > 0]
write_csv(naxos_diet_matrix, 'output_data/naxos_diet_matrix.csv', col_names=TRUE)

```

**Site by OTU matrix (Naxos): Prey**

```{r, include=TRUE, echo=FALSE}

kable(naxos_prey_matrix) %>% 
  kable_styling(bootstrap_options = c('striped', 'condensed'),
                full_width = FALSE,
                position = 'left')

```

<br>

**Site by OTU matrix (Naxos): Diet**

```{r, include=TRUE, echo=FALSE}

kable(naxos_diet_matrix) %>% 
  kable_styling(bootstrap_options = c('striped', 'condensed'),
                full_width = FALSE,
                position = 'left')

```
