---
title: "Data Wrangling"
author: "Kate Culhane"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(include=FALSE)
```

<br>

### Pre-analysis manipulation of datasets

CSVs created for use in analyses, stored in "output_data".

<br>

**Prey data:**

- *prey.csv*: prey data from both sticky and pitfall traps in long format (1 row per OTU per sticky/pitfall trap pair)
- *naxos_prey.csv*: prey data from both sticky and pitfall traps, for Naxos only in long format (1 row per OTU per sticky/pitfall trap pair)

<br>

**Diet data:**

- *diet.csv*: diet data from field samples in long format (1 row per OTU per lizard)
- *naxos_diet.csv*: diet data from field samples, for Naxos onlyin long format (1 row per OTU per lizard)

<br>

**Site by OTU matrices:**

- *naxos_prey_matrix.csv*: Site by OTU matrix for prey data, for Naxos data only
- *naxos_diet_matrix.csv*: Site by OTU matrix for diet data, for Naxos data only

<br>

```{r packages and data}

# Load packages
library(kableExtra)
library(tidyverse)

# Diet data
field_diet <- read_csv('raw_data/diet_field.csv') # from field samples

# Prey data
pitfall <- read_csv('raw_data/prey_pitfall.csv') %>% # from pitfall traps
  mutate(trap_id = paste0(site, '-', trap),
         type = 'pitfall')
sticky <- read_csv('raw_data/prey_sticky.csv') %>% # from sticky traps
  mutate(trap_id = paste0(site, '-', trap),
         type = 'sticky')
prey_meta <- read_csv('raw_data/prey_meta.csv') # trap metadata

# Naxos site data
naxos_meta <- read_csv('raw_data/naxos_meta.csv') # metadata for Naxos sites
naxos_veg <- read_csv('raw_data/naxos_veg.csv') # veg quadrat data for Naxos sites

```

```{r}

# Prey data

# Combine pitfall and sticky data by trap
prey <- bind_rows(pitfall, sticky) %>% 
  select(-c(date_analyzed, condition, comments, type)) %>% 
  group_by(island, site, trap, trap_id) %>% 
  summarize_all(sum, na.rm=TRUE) %>% 
  ungroup() %>% 
  pivot_longer(coleoptera:homoptera, names_to = 'otu', values_to = 'n') %>% 
  filter(n != 0)
write_csv(prey, 'output_data/prey.csv', col_names=TRUE)

# Naxos data only
naxos_prey <- prey %>% 
  filter(island == 'Naxos') %>% 
  full_join(naxos_meta[c(1:4)]) %>% 
  select(site, use, veg, trap, trap_id, otu, n)
write_csv(naxos_prey, 'output_data/naxos_prey.csv', col_names=TRUE)

### JOIN prey AND naxos_prey WITH METADATA ABOUT BUGS (SIZE, HARDNESS, TAXON, TYPE, ETC.)

```

```{r}

# Diet data

# All data
diet <- field_diet %>% 
  select(-c(date, comments)) %>% 
  pivot_longer(blattodea:salt, names_to = 'otu', values_to = 'n') %>% 
  filter(n != 0)
write_csv(diet, 'output_data/diet.csv', col_names=TRUE)

# Naxos data only
naxos_diet <- diet %>% 
  filter(island == 'Naxos') %>% 
  full_join(naxos_meta[c(1:4)]) %>% 
  select(site, use, veg, lizard_id, otu, n)
write_csv(naxos_diet, 'output_data/naxos_diet.csv', col_names=TRUE)

```

```{r}

# Site by OTU matrices

# Prey
naxos_prey_matrix <- naxos_prey %>%
  select(-c(trap, trap_id)) %>% 
  group_by(site, use, veg) %>% 
  summarize_all(sum) %>% 
  arrange(use, veg)
naxos_prey_matrix <- naxos_prey_matrix[, colSums(naxos_prey_matrix != 0) > 0]
write_csv(naxos_prey_matrix, 'output_data/naxos_prey_matrix.csv', col_names=TRUE)

# Diet
naxos_diet_matrix <- naxos_diet %>% 
  select(-c(lizard_id, comments)) %>% 
  group_by(site, use, veg) %>% 
  summarize_all(sum) %>% 
  arrange(use, veg)
naxos_diet_matrix <- naxos_diet_matrix[, colSums(naxos_diet_matrix != 0) > 0]
write_csv(naxos_diet_matrix, 'output_data/naxos_diet_matrix.csv', col_names=TRUE)

```

**Site by OTU matrix (Naxos): Prey**

```{r, include=TRUE, echo=FALSE}

kable(naxos_prey_matrix) %>% 
  kable_styling(bootstrap_options = c('striped', 'condensed'),
                full_width = FALSE,
                position = 'left')

```

<br>

**Site by OTU matrix (Naxos): Diet**

```{r, include=TRUE, echo=FALSE}

kable(naxos_diet_matrix) %>% 
  kable_styling(bootstrap_options = c('striped', 'condensed'),
                full_width = FALSE,
                position = 'left')

```
