---
title: "Naxos - Exploratory Analyses"
author: "Kate Culhane"
output:
  html_document:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = '..')
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width=7, fig.height=5)
```

```{r include=FALSE}

# Load necessary packages

library(vegan)
library(tidyverse)

# Load necessary data

diet <- read_csv('raw_data/diet_field.csv') %>% # diet data from field
  filter(island == 'Naxos')

inverts_meta <- read_csv('raw_data/inverts_meta.csv') %>% # metadata for invert community field collection
  filter(island == 'Naxos')

pitfall <- read_csv('raw_data/inverts_pitfall.csv') %>% # invert community data from pitfall traps
  filter(island == 'Naxos') %>% 
  mutate(trap_id = paste0(site, '-', trap))

sticky <- read_csv('raw_data/inverts_sticky.csv') %>% # invert community data from sticky traps
  filter(island == 'Naxos') %>% 
  mutate(trap_id = paste0(site, '-', trap))

site_meta <- read_csv('raw_data/site_meta.csv') # metadata for Naxos sites

site_veg <- read_csv('raw_data/site_veg.csv') # vegetation quadrat data for Naxos sites

```

## Guiding Questions

**1: Understanding variation in the invertebrate community on Naxos**

- Is there variation in the invertebrate community across nine different sites on Naxos?
- Is that variation related to vegetation characteristics? (% cover of different proportions of vegetation)
- Is that variation related to human land-use metrics? (presence/absence of goats or walls)

**2: Understanding diet of *P. erhardii* on Naxos**

- Is there variation in diet of the lizards across the nine sites?
- Is any variation in diet related to the variation in insect availability? (Are the lizards generalists or specialists? Are they choosy? Are they eating just in proportion to what they're seeing?)

## Analyses to do

1. NMDS for prey communities
2. NMDS for lizard diet
3. Adonis between site types

## Prey data

### Pitfall

```{r}

### NMDS prep

# Change df to matrix
pitfall_m <- as.matrix(pitfall[-c(1:4,29:31)])
rownames(pitfall_m) <- pitfall$trap_id

# Choose a distance matrix
rankindex(pitfall$site,
          wisconsin(pitfall_m),
          method = 'spearman')

```



## Diet data

```{r, include=FALSE}

# Update class of the data frame to "matrix"

diet_m <- as.matrix(diet[-c(1:4,44)]) # remove columns that aren't species
rownames(diet_m) <- diet$lizard_id # name the columns using site names

# Choose a distance metric

rankindex(diet$site,
          wisconsin(matrix),
          indicies = c('bray', 'euclidean', 'manhattan', 'horn'),
          method = 'spearman')

```

```{r}

# Ordination

ord <- metaMDS(matrix, distance = 'bray', k = 2)
ord$stress
stressplot(ord)

# Plot

par(mar = c(4,4,1,1))

plot(ord, display = 'sites', type = 'n')

points(ord, display = 'sites')

ordiellipse(ord, groups = diet$site)

text(ord, display = 'species')

```


```{r eval=FALSE}

rmarkdown::render('code/naxos_exploratory.Rmd', output_file = '../docs/naxos_exploratory.html')

```
